---
title: "STAT 3355 HW1"
author: "Adyan Rahman"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

(a)
```{r}
x <- 8 + 9 - (7/3^(0.3))
round (x, digits <- 2)

```

(b)
```{r}
y <- sqrt((15 + 16) / (14 + 12))
round(log2(y), digits <- 2)
```

(c)
```{r}
x <- (11 + sin(pi / 4))
y <- (factorial(3) + abs(-11))
z <- x / y
round(z ^ 2, digits <- 2)
```

(d)
```{r}
round(6 + 5 - (4/(3 ^ 2)), digits <- 2)
```

(e)
```{r}
x <- 14 + 13
y <- 12 + 11
round(exp(sqrt(x / y)), digits <- 2)
```

(f)
```{r}
x <- 11 + factorial(12)
y <- factorial(13) + 14
round((x / y) ^ 2, digits <- 2)
```

# Problem 2

(a) Putting data into data vector RF
```{r}
RF <- c(2.60, 3.05, 3.74, 3.48, 5.49, 4.25, 2.57, 
        2.18, 3.14, 4.82, 3.28, 3.01)
```

(b) Renaming RF to mon
```{r}
mon <- RF
```

(c) Average rainfall during the year
```{r}
round(mean(mon), digits <- 2)
```

(d) Finding months with minimum and maximum rainfall
```{r}
which.min(mon)
which.max(mon)
```

# Problem 3

(a) Putting data in vector HF
```{r}
H2 <- c(2700, 2600, 3050, 2900, 3000, 2500, 2600, 3000, 2800, 3200, 2800, 3400)
```

(b) Renaming vector to mon
```{r}
mon <- H2
```

(c) Total Hummers sold in 2002
```{r}
sum(mon)
```

(d) Finding greatest increase in decrease
```{r}
# Greatest increase within a month
monthly_difference <- diff(mon)
monthly_increase <- which.max(monthly_difference)
greatest_increase <- monthly_increase + 1
print(greatest_increase)

# Greatest decrease within a month
monthly_decrease <- which.min(monthly_difference)
greatest_decrease <- monthly_decrease + 1
print(greatest_decrease)
```

# Problem 4

(a)
```{r}
x <- c(1, -2, 3, -4, 5, 100)
y <- x * -1
y[ y > 0]
```

(b)
```{r}
# create a sequence from 1 to 50
z <- seq(1 : 50)

# test whether an observation is even
even <- z %% 2 == 0

# subset z by the test above
z <- z [even]
```

(c)
```{r}
mean <- function(x) {
  sum(x) / length(x)
}
```

# Problem 5

Find perfect squares
```{r}
PrintSquare <- function() {
  for (num in 1:1000) { 
    if (sqrt(num) == as.integer(sqrt(num))) {
      print(num)
    }
  }
}

PrintSquare()
```
 \newpage

# Problem 6

Find twin primes
```{r}
prime <- function(num) {
  if (num <= 1) {
    return(FALSE)
  }
  for (i in 2:(sqrt(num))) {
    if (num %% i == 0) {
      return(FALSE)
    }
  }
  return(TRUE)
}


twin_primes <- function(n) {
  twin_primes_num <- 0
  
  
  for (num in 1 : (n - 2)) {
    if (prime(num) && prime(num + 2)) {
      twin_primes_num <- twin_primes_num + 1
    }
  }
  
  return(twin_primes_num)
}

#n <- readline(Prompt <- "Enter a number")
#n <- as.numeric(n)
#print(twin_primes(n))

# Example Output
n <- 50
print(twin_primes(n))
```















